Reduced diversity and increased specialization of Alaska salmon fishermen:

### 2017-02-22 salmon/paper/salmon_paper_v3.Rmd
Revert "Revert "local change/""
This reverts commit 240797038b8aada993df17a517b92bf49f992ec8.

---
title: 'Reduced diversity and increased specialization of Alaska salmon fishermen:
  effects on variation in individual revenues'
author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, Ben, etc."
date: '2017-02-08'
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
bibliography: CatchPortfolios_salmon.bib
---

@@

On a large scale, individual fishers may experience positive outcomes from diversifying their catches; examples include reduced year-to-year variability in annual revenue and increases in average revenue. Fishers can increase their species diversity by owning a permit that allows multiple species to be targeted, or by owning multiple permits. External forces such as management regulations (limited entry,  catch shares), the number of permits available, or market forces (permit or fish prices) may disincentive individuals to hold multiple permits. Thus, individuals holding a single permit may only be able to increase species diversity within the confines of their permit (longer fishing seasons, broader spatial area). Using a long-term dataset of individual salmon fishers in Alaska, we develop new models for understanding how the effects of diversification have changed through time. These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production, market forces, EVOS?).

@@

Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski_income_2013; @anderson_benefits]. Fisheries in Alaska have been studied extensively, because they are one of the most productive seafood regions in the world (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies. [@anderson_benefits] defined strategies as a permit or collection of permits held by single individuals. While adding new permits allows fishers to increase species diversity, few individual permits allow individuals to harvest more than one species (examples of single-species permits in the Gulf of Alaska include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber). Some exceptions to these single species permits include participants in federally managed groundfish fisheries and participants in state managed salmon fisheries.

Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits] (figure XX). For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig. A1), median revenue is increasing (across individuals, Fig. A2), and landed species diversity declining over time (averaged across individuals, Fig. A6). Across different salmon fisheries -- other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]) -- there appears to be little correlation between diversity and revenue (Fig. A4). In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV of revenue, but not both [@purdy_financial_1997], and the same result may be true for salmon fishers. Further, these impacts of diversification are variable over time.

Five species of salmon are harvested in Alaska waters (Chinook [*Oncorhynchus tshawytscha*], chum [*O. keta*], pink [*O. gorbuscha*], sockeye [*O. nerka*], coho [*O. kisutch*]), and the ability of individual salmon fishers to diversify across species is in part a function of geography and fishing gears. For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), and set gillnets (12 areas). Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska, respectively). While some of these combinations of gears and areas allow fishers to diversify, there are many examples of permits where the catch is dominated by a single species. Perhaps the best examples include fisheries that target sockeye salmon (e.g. Bristol Bay set and drift gillnet fisheries). For these fisheries, specialization is associated with higher revenue and reduced CV in annual revenue [@anderson_benefits].

A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals. The modern era of Alaska’s large-scale hatchery program began in the late 1970s (Stopha 2015, Rich will get a reference), with substantial increases in hatchery production during the 1980s following low returns [@mahnken_historical_1998; @leber_stock_2008-1]. Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn_review_2001], though these programs may also have negative impacts on wild production [@wertheimer_relationship_2005; @ruggerone_productivity_2015]. The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn_review_2001; @hilborn_hatcheries_1992; @wertheimer_comment:_2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber_stock_2008-1] or economic impacts [@boyce_alaska_1993], rather than to variability experienced by individual fishermen.

[Jordan: Maybe a paragraph or notes about the importance of the role that salmon are serving in buffering or stabilizing fishing revenues in the state and how subsequently, understanding the responses of salmon to climate dynamics is especially critical (Mueter et al. 2002, and maybe Schindler et al., 2010)]

In addition to the changing species composition of fish available to be harvested, the benefits (or costs) of diversification are affected by a number of exogenous drivers, including the influx of farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends]. [Break out as own paragraph+add more examples?] Following price crashes in the 1990s and early 2000s (cite CFEC), many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig. A3, A12). In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies). Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives for diversification.

Tradeoffs between specialization and diversification may also be found in case studies of terrestrial agriculture. For small scale farmers, increasing crop diversity may increase revenue and stabilize year-to-year income variability [@di_falco_crop_2003]. At larger scales or in the developed world, however, the incentives for diversification may be different. In the United States and Canada for example, crop diversity has declined since the late-1970s [@aguilar_crop_2015; @bradshaw_farm-level_2004]; some of the factors responsible include changing national policies, technology, and increased demand from other parts of the world (e.g. increased U.S. production of soybeans to meet demand from Asia). Shifts to from diverse of high value cash-crops to specializing on high-volume low-value crops such as cereals may have analogs for salmon fishermen in low value species with rapid life histories (pink and chum salmon).

The objective of our analysis is to better understand the effects of diversification on single-permit salmon holders by extending the methods used by [@anderson_benefits] to include time-varying effects. These time-varying effects will allow us to compare temporal trends across regions, identify periods when having diverse catch portfolio was most beneficial, and allow us to evaluate support for effects of longer term drivers in the region.

@@

We obtained fisheries landings and revenue data for all iundividual permit holders in Alaska from 1985-2014 from the Commercial Fisheries Entry Commission (CFEC). Each permit holder has a unique identifier that allows us to track quantities including the species caught and revenue received. We adjusted for inflation by converting all revenues to 2009 USD. Because of our focus on salmon fishers, we were only interested in single permit fisheries, and we adopted a simplified version of the filtering steps described in [@anderson_benefits]. After aggregating individual trip-level catch and revenue to annual data, we retained permit holders who fished only one of the six salmon permits included in a year (Table 1, Fig. A5). Like [@anderson_benefits] we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies (e.g. changing from a seine fishery in Kodiak to a gillnet fishery in Bristol Bay), we only included pairs of years for a given person where they did not change strategies (i.e., didn’t change fisheries or add permits, etc.).

@@

### As hatchery programs have increased production, have individual salmon fishermen experienced reduced variability?

Hatchery release trends for some regions included in our data have been summarized previously (e.g. [@ward_evaluating] summarized the trends for Prince William Sound). Release patterns have been dominated by pink and chum salmon ([@stopha2016]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014). Though our analysis with CFEC data doesn't separate hatchery versus wild fish, as a contrast we can make comparisons across regions and species to examine potential effects of hatchery fish on variability. For example, we can compare the variation in the largely hatchery pink salmon returns (earnings and weight) from Prince William Sound to the largely wild pink salmon returns from Southeast Alaska. Similarly, we can compare the variability in earnings from species dominated by hatchery returns (pink, chum) to those with less hatchery influence (Chinook, sockeye). To make these comparisons, we aggregated the earnings and landings by species and permit, then calculated 3-year rolling CVs. As a second comparison, we examined whether long-term participants in any of the fisheries had experienced a change in variability. Similar to [@kasperski_income_2013] who calculated CVs of vessels before and after significant management events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014).

### Analysis 3: [Core of paper] Does effect of diversification change over time?

[@anderson_benefits] implemented a hierarchical Bayesian model that allowed group level predictors and covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers. This model also included random intercepts by strategy and year; these terms allowed all permit holders in a given year to experience common fluctuations (external drivers including price changes, fluctuations in fish population abundance, or environmental disasters such as EVOS).

To ask how benefits of diversification may have changed over time, we extend the model of [@anderson_benefits] to also include time-varying random effects in the slope coefficients of diversification. We define the variables $\Delta { S }_{ t }=\log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=\log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years. The model for the expected log revenue of individual *i* in year *t*, $$\widehat { \log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ i,t }+\log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1). We model the time-varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim \mathrm{Normal}\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$. The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species. Our model extends that of [@anderson_benefits] to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification. We assume this relationship is linear, and the coefficients ${ B }_{ 1,j,t }$ and ${ B }_{ 2,j,t }$ allow this to vary by permit and through time. The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent of alterations to catch composition). We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim \mathrm{Normal}\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$. Like [@anderson_benefits] we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time. Finally, the term $\log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits]. 
  Like [@anderson_benefits], we also included all of the above covariates in the variance model. We assumed the residuals from the mean model described above to be normally distributed, $$\log\left( { R }_{ i,t } \right) \sim \mathrm{Normal}\left( \widehat { \log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$. While conventional hierarchical linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models (CITE). Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { \log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$. Unlike the linear model for the mean, we did not include an offset term. Further, we constrained the variance intercepts ${ G }_{ 0,j }$ to be temporally constant by strategy, instead of the dynamic nature of the intercepts in the mean model. Like the mean model, the random effects were constrained to be autoregressive, ${ G }_{ 1,j,t }\sim \mathrm{Normal}\left( { G }_{ 1,j,t-1 },{ \sigma  }_{ { G }_{ 1 } } \right)$, and covariates associated with days fished and species diversity were included as predictors.

[Rich's note: This and delta D might be tricky for pink salmon stocks, which alternate between strong even or odd years (depends on the area) and may dictate fishery participation. For these, could we, should we, use log(St/St-2)??]

All done in Stan [@stan_development_team_stan_2015] using the NUTS algorithm [@hoffman_no-u-turn_2014].

Specific things that could be looked at: 
- EVOS (closure for Kodiak)  
- PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options
- sockeye marketing 1983: [Jordan: So the Copper River salmon marketing ramped up in about 1983. I wonder if we could see a steady increase in value or movement towards sockeye in the early years of the data set]

Does diversification exist within permits and if so, do we see similar advantages of diversification on revenue + CV?

PWS chum and sockeye (increasing hatchery % over time) vs. Bristol Bay (all wild).

CV of variability for high value species (sockeye, Chinook) vs. low value species (pink and chum). Coho fall somewhere in the middle. In other words, how have choices of species to produce by hatcheries affected the CV of revenue for individual permits? % hatchery fish of the harvest could be used as a continuous variable in the model.

Similar to above, do life-history characteristics influence the variability of revenue? For example, sockeye and chum have substantially more complicated life-history traits (1-5 ocean years) than pink salmon (1 ocean year only). For high-volume hatcheries such as pink salmon: do annual returns affect price and revenue the following year more than for high-value fisheries?  
## Results

Our summaries of changing variability (aggregate for each fishery, across individuals, or within individuals) was not consistent with the hypothesis that variability in revenue or landings has decreased over time (Fig. A9-A10). For pink and chum salmon, fisheries in PWS generally experience higher year-to-year variability compared to SEAK. In looking at the changing CV of individuals, we see a slight increase in the CV of revenue for people fishing in the recent decade who were also fishing 1985-1995 (with the exception being S01A).

@@

- For small scale farmers, economists have often wondered why farmers don't diversify more. Diversification can be profitable, but in the short term risky or expensive. Bigger than cost seems to be the risk [@karlan_agricultural_2012]

@@

- One of the outputs that can be calculated from the raw data is variability of revenue over time (e.g. year to year, at the level of individual & permit). This gets at more of the social impacts of changing opportunities, but long term trends in rising income variability have been documented in the US (Hacker 2006). Understanding trends in revenue of individuals whose incomes are dependent on natural resources are interesting because they tend to be more variable than that of the larger population [@mishra_income_2002]. Are there specific attributes that correlate with rising variability (for example the CBO has suggested variability in year to year income is negatively correlated with income [@cbo_recent_2008])

### Robustness of strategies to changing environments [stash in appendix?]

One of the challenges in modeling the effects of species diversification for individual fishers is that between two years, both the species composition of the catch and the species composition of the fish available to be caught change (making it hard to disentangle the selectivity of fishermen versus portfolio of fish available). To isolate the effects of a changing environment from changing strategies, we used a subset of the data from fishery permits where discrete strategies were emergent. For details see Appendix.

Using purse seine permit holders in Prince William Sound as an example, we first calculated the species composition (weighted by revenue) for each active permit holder in each year. For each pair of years that an individual fished, we calculated the squared Euclidian distance between the species compositions of those years, $$\Delta { p }_{ ind,t }=\sum _{ sp=1 }^{ 5 }{ { \left| { p }_{ ind,sp,t }-{ p }_{ ind,sp,t-1 } \right|  }^{ 2 } } $$.

Larger values of $\Delta { p }_{ ind,t }$ indicate greater year-to-year discrepancy in species composition. To focus on individuals who had more constant strategies in adjacent years, we ommitted data corresponding to the upper third of this distribution (retaining $\Delta { p }_{ ind,t }$ < 0.334). Next, we used the species composition from each individual and pair of years to cluster individuals into strategies. Clustering was performed using the partitioning around medoids (pam) algorithm in R with an estimated number of clusters (using 'pamk' in fpc library), and selected 2 groups of individual strategies: pink salmon specialization (medoid proportion of pink salmon = 0.95), and a strategy focused on pink salmon and chum (medoid proportions of 0.64 and 0.31 respectively). For each group, we then calculated the mean and CV of revenue; we calculated the log difference of revenue for each individual, $\Delta { R }_{ i }=\log\left( { R }_{ i,t+1 }/{ R }_{ i,t } \right)$ and then calculated the CV as the standard deviation over people, $sd(\Delta { R }_{ i=1:n })$.

@@

```{r, echo=FALSE, fig.pos="placeHere", fig.cap="CV and mean annual revenue by salmon permit, 1975-2014. Average participation (unique fishers) given by circle size, and effective species diversity is shown with color shading.", fig.height=3.2, eval=FALSE}
dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")

cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
people = group_by(cfec_all, p_fshy, year) %>% 
  summarize(n = length(unique(p_holder))) %>% 
  group_by(p_fshy) %>% summarize(n = mean(n,na.rm=T))
names(people)[which(names(people)=="p_fshy")] = "strategy_permit"
people$strategy_permit = paste0(substr(people$strategy_permit,1,1), substr(people$strategy_permit,3,5))

dat.annual = filter(dat.annual, npermit==1 & substr(strategy_permit,1,1)=="S") %>%
  group_by(strategy_permit) %>% 
  mutate(nPeople = length(unique(p_holder))) %>% 
  filter(nPeople > 200) %>% 
  dplyr::select(-nPeople) %>% 
  group_by(strategy_permit, year) %>% 
  summarize(meanrev = mean(revenue), meandiv = mean(specDiv)) %>% 
  arrange(strategy_permit, year) %>% 
  group_by(strategy_permit) %>% 
  summarize(mean_rev = mean(meanrev,na.rm=T), cv_rev = sd(diff(log(meanrev)),na.rm=T),
    diversity = mean(meandiv,na.rm=T)) %>% as.data.frame

>>>>

---
title: 'Reduced diversity and increased specialization of Alaska salmon fishermen:
  effects on variation in individual revenues'
author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, Ben, etc."
date: ''
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
bibliography: CatchPortfolios_salmon.bib
---

@@

On a large scale, individual fishers may experience positive outcomes from diversifying their catches; examples include reduced year-to-year variability in annual revenue and increases in average revenue. Fishers can increase their species diversity by owning a permit that allows multiple species to be targeted, or by owning multiple permits. External forces such as management regulations (limited entry,  catch shares), the number of permits available, or market forces (permit or fish prices) may disincentive individuals to hold multiple permits. Thus, individuals holding a single permit may only be able to increase species diversity within the confines of their permit (longer fishing seasons, broader spatial area). Using a long-term dataset of individual salmon fishers in Alaska, we develop new models for understanding how the effects of diversification have changed through time. These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production, market forces, EVOS).

@@

Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski2013; @anderson_benefits]. Fisheries in Alaska have been studied extensively, because they are one of the most productive seafood regions in the world (both by landed weight, and revenue [@officeofscienceandtechnology2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies. [@anderson_benefits] defined strategies as a permit or collection of permits held by single individuals. While adding new permits allows fishers to increase species diversity, few individual permits allow individuals to harvest more than one species (examples of single-species permits in the Gulf of Alaska include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber). Some exceptions to these single species permits include participants in federally managed groundfish fisheries and participants in state managed salmon fisheries.

Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits]. For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig. A1), median revenue is increasing (across individuals, Fig. A2), and landed species diversity declining over time (averaged across individuals, Fig. A6). Across different salmon fisheries -- other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]) -- there appears to be little correlation between diversity and revenue (Fig. 2). In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV of revenue, but not both [@purdy1997], and the same result may be true for salmon fishers. Further, these impacts of diversification are variable over time.

Five species of salmon are harvested in Alaska waters (Chinook [*Oncorhynchus tshawytscha*], chum [*O. keta*], pink [*O. gorbuscha*], sockeye [*O. nerka*], coho [*O. kisutch*]), and the ability of individual salmon fishers to diversify across species is in part a function of geography and fishing gears. For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), and set gillnets (12 areas). Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska, respectively). While some of these combinations of gears and areas allow fishers to diversify, there are many examples of permits where the catch is dominated by a single species. Perhaps the best examples include fisheries that target sockeye salmon (e.g. Bristol Bay set and drift gillnet fisheries). For these fisheries, specialization is associated with higher revenue and reduced CV in year-to-year revenue [@anderson_benefits].

A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals. The modern era of Alaska’s large-scale hatchery program began in the late 1970s [@stopha2016], with substantial increases in hatchery production during the 1980s following low returns [@mahnken1998; @leber2008]. Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn2001a], though these programs may also have negative impacts on wild production [@wertheimer2005; @ruggerone2015]. The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn2001a; @hilborn1992a; @wertheimer2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber2008] or economic impacts [@boyce1993], rather than to variability experienced by individual fishermen. Understanding the role salmon have in buffering or stabilizing fishing revenues is critical given the responses of salmon to variable environments [@mueter2002; @schindler2010].

In addition to the changing species composition of fish available to be harvested, the benefits or costs of diversification are affected by a number of exogenous drivers, including competition from markets producing farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends]. Following price declines in the 1990s and early 2000s [@adfg2007], many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig. A3, A12). In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies). Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives or opportunities for diversification. 
Tradeoffs between specialization and diversification may also be found in case studies of terrestrial agriculture. For small scale farmers, increasing crop diversity may increase revenue and stabilize year-to-year income variability [@difalco2003]. At larger scales or in the developed world, however, the incentives for diversification may be different. In the United States and Canada for example, crop diversity has declined since the late-1970s [@aguilar2015; @bradshaw2004]; some of the factors responsible include changing national policies, technology, and increased demand from other parts of the world (e.g. increased U.S. production of soybeans to meet demand from Asia). Shifts from more diverse high value crops to specializing on high-volume low-value crops such as cereals may have analogs for salmon fishermen in low value species with rapid life histories (pink and chum salmon).

The objectives of our analysis are to (1) evaluate long term trends in stability for fleets and individuals to evaluate whether hatchery production has increased stability, and (2) to better understand the effects of diversification on single-permit salmon holders by extending the time series models of individual revenues used by [@anderson_benefits] to include time-varying effects. These time-varying effects of diversification will allow us to compare temporal trends across regions and identify periods when having diverse catch portfolio was most beneficial.

@@

We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985-2014 from the Commercial Fisheries Entry Commission (CFEC). Reported gross earnings (revenue) associated with each fish ticket were adjusted for inflation by converting all revenues too 2009 USD. Because our modeling focus involves estimating time varying effects of diversification, we limited our analysis to salmon fisheries with species diversity > 1 (omitting fisheries such as the Bristol Bay gillnet fisheries that specialize on sockeye salmon). To better understand the influence of changing hatchery production, we also restricted fisheries to those that have been influenced by hatchery salmon production.

@@

### Has increased hatchery production stabilized Alaskan salmon fisheries?

Hatchery release trends for some regions included in our data have been summarized previously (e.g. [@ward_evaluating] summarized the trends for Prince William Sound). Release patterns have been dominated by pink and chum salmon ([@stopha2016], Fig. 3) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014). In addition to examining temporal trends, we can use the spatial contrast between areas to quantify potential changes resulting from increased hatchery production (Prince William Sound, Kodiak, Southeast Alaska).

To examine long term stability in salmon fisheries, we summarized the trends in landings (total pounds) and revenue (2009 USD) at two levels. First, we calculated annual summaries of catch and revenue at the fishery level (annual values summed across all permit holders) and calculated 3-year running coefficients of variation (CV) by fishery. Examining fishery-level summaries alone may be problematic, because increases in individual variability may have a stabilizing effect at larger scales (communities, fisheries). As a second metric of variability, we calculated the 3-year running CV for every unique permit holder in each fishery, and averaged those individual CVs across permit holders. To contrast patterns across species, we repeated these calculations, computing the annual CV of revenue and prices separately for each species and fishery combination.

### Have effects of diversification changed over time for salmon fishers?

Because we were only interested in single permit fisheries, we applied a simplified version of the CFEC data filtering steps described in [@anderson_benefits]. Individual trip-level catch and revenue statistics were aggregated to annual data, and we retained individual fishers who only participated in one fishery in a year (Table 1). Annual revenues were log-transformed to remove non-stationary individuals; to avoid complications in modeling individual switching permits, we only included pairs of years for a given person where they did not change permits (e.g. didn’t change fisheries between two years). [@anderson_benefits] implemented a hierarchical Bayesian time series model that incorporated group level predictors and individual covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers. This model also included random intercepts by fishery and year, allowing all permit holders in a given year to experience common fluctuations (external drivers including price changes, fluctuations in fish population abundance, or environmental disasters such as EVOS).

To ask how benefits of diversification may have changed over time for salmon fisheries in Alaska, we extend the model of [@anderson_benefits] to include time-varying random effects in the slope coefficients of diversification. We define the variables $\Delta { S }_{ t }=\log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=\log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years. The model for the expected log revenue of individual *i* in year *t*, $$\widehat { \log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ i,t }+\log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1). We model the time-varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim \mathrm{Normal}\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$. The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species. Our model extends that of [@anderson_benefits] to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification. We assume this relationship is linear, and the coefficients ${ B }_{ 1,j,t }$ and ${ B }_{ 2,j,t }$ allow this to vary by permit and through time. The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent of alterations to catch composition). We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim \mathrm{Normal}\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$. Like [@anderson_benefits] we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time. Finally, the term $\log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits].

Like [@anderson_benefits], we also included all of the above covariates in the variance model. We assumed the residuals from the mean model described above to be normally distributed, $$\log\left( { R }_{ i,t } \right) \sim \mathrm{Normal}\left( \widehat { \log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$. While conventional hierarchical linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models (CITE). Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { \log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$. Unlike the linear model for the mean, we did not include an offset term. Further, we constrained the variance intercepts ${ G }_{ 0,j }$ to be temporally constant by strategy, instead of the dynamic nature of the intercepts in the mean model. Like the mean model, the random effects were constrained to be autoregressive, ${ G }_{ 1,j,t }\sim \mathrm{Normal}\left( { G }_{ 1,j,t-1 },{ \sigma  }_{ { G }_{ 1 } } \right)$, and covariates associated with days fished and species diversity were included as predictors.

All estimation was done in a Bayesian framework using R [@rcoreteam2016] and the package 'rstan' implementing using the NUTS algorithm [@hoffman2014].

Specific things that could be looked at: 
[Rich: maybe look at PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options. EW: pink harvest generally falls from 89/90 to early 90s. It's a little of a chicken and egg problem too - because by the early 90s hatchery production was 75% of the total pink catch and also declined a little bit. From the CV perspective, the pink crash may be synchronizing -- stability was increased because there were several back to back bad years]
[Jordan: So the Copper River salmon marketing ramped up in about 1983. I wonder if we could see a steady increase in value or movement towards sockeye in the early years of the data set. EW: Fig 3 largely shows the pattern in those early years -- it ramps up in PWS, but is less steep than for say chum]

## Results

Our summaries of changing variability (aggregate for each fishery, or averaged across individuals) are not consistent with the hypothesis that variability in revenue or landings has decreased over time (Fig. A9-A10). For pink and chum salmon, fisheries in PWS generally experience higher year-to-year variability compared to SEAK. In looking at the changing CV of individuals, we see a slight increase in the CV of revenue for people fishing in the recent decade who were also fishing 1985-1995 (with the exception being S01A).

@@

- For small scale farmers, economists have often wondered why farmers don't diversify more. Diversification can be profitable, but in the short term risky or expensive. Bigger than cost seems to be the risk [@karlan2012]

@@

- One of the outputs that can be calculated from the raw data is variability of revenue over time (e.g. year to year, at the level of individual & permit). This gets at more of the social impacts of changing opportunities, but long term trends in rising income variability have been documented in the US (Hacker 2006). Understanding trends in revenue of individuals whose incomes are dependent on natural resources are interesting because they tend to be more variable than that of the larger population [@mishra2002]. Are there specific attributes that correlate with rising variability (for example the CBO has suggested variability in year to year income is negatively correlated with income [@cbo2008])

@@

```{r fig1, echo=FALSE, fig.pos="placeHere", fig.cap="Trends in participation, landed catch, mean annual revenue, and effective species diversity for eight selected salmon fisheries. [EW notes: weight per person increasing more than revenue. Species diversity for most fisheries declining]"}
# trends in participation, 
cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
people = group_by(cfec_all, p_fshy, year) %>% 
  summarize(n = length(unique(p_holder))) %>% 
  filter(substr(p_fshy,1,1)=="S")
people$p_fshy = paste0(substr(people$p_fshy,1,1), substr(people$p_fshy,3,5))
names(people)[which(names(people)=="p_fshy")]="Permit"

g1 = ggplot(people[people$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(n), color = Permit)) + geom_line() + ylab("Log10 unique people") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

# trends in landed weight - metric tons
weight = group_by(cfec_all, p_fshy, year) %>% 
  summarize(w = sum(g_pounds,na.rm=T) * 0.000453592) %>% 
  filter(substr(p_fshy,1,1)=="S")
weight$p_fshy = paste0(substr(weight$p_fshy,1,1), substr(weight$p_fshy,3,5))
names(weight)[which(names(weight)=="p_fshy")]="Permit"

weight = left_join(weight, people)
g2 = ggplot(weight[weight$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(w/n), color = Permit)) + geom_line() + ylab("Log10 tons per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

# trends in revenue
dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")
dat.annual = group_by(dat.annual[which(dat.annual$strategy_permit %in% c("S01A", "S01E", "S01K", "S01M", "S03A", "S03E", "S03T", "S05B")),], year, strategy_permit) %>% 
  summarize(meanrev = mean(revenue), meandiv = mean(specDiv))
names(dat.annual)[which(names(dat.annual)=="strategy_permit")]="Permit"
dat.annual = left_join(dat.annual, weight)
